---
title: "C02 Goodies"
author: "David Jackson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(duckdb)
library(janitor)
library(scales)
```

#### Read CVS files 

```{r}
rm(list=ls())
passenger <- read_csv("CarbonEmissions/carbon-new-passenger-vehicles.csv") %>% clean_names()
```
```{r}
effciency <- read_csv("CarbonEmissions/fuel-efficiency-new-vehicles.csv") %>% clean_names()
```

#### Create DuckDB tables

```{r}
con <- dbConnect(duckdb::duckdb(), dbdir = ":memory:")
dbWriteTable(con, "passenger", passenger)
dbWriteTable(con, "effciency", effciency)
# dbDisconnect(con)
```

```{r}
dbGetQuery(con,"SELECT * 
           FROM passenger 
           LIMIT 1")
```
```{r}
dbGetQuery(con,"SELECT * 
           FROM effciency 
           LIMIT 1")
```

## Join Tables on Code and Year

```{r}
fuel_efficiency <- dbGetQuery(con,"
           SELECT co2.entity as country,co2.year,co2.co2_per_km,
           fuel_efficiency_per_km
           FROM passenger co2
           INNER JOIN effciency eff
           ON co2.code = eff.code AND co2.year = eff.year;")
```
```{r}
dbGetQuery("DESCRIBE SELECT * FROM fuel_efficiency;")
```

Sure! Here's some sample SQL code to calculate summary statistics for the CO2 emissions per km and fuel efficiency per km for each country and year:

```{r}
dbGetQuery(con,"
           SELECT
          co2.entity AS country,co2.year,
          AVG(co2.co2_per_km) AS avg_co2_per_km,
          STDDEV(co2.co2_per_km) AS std_dev_co2_per_km,
          AVG(eff.fuel_efficiency_per_km) AS avg_fuel_efficiency_per_km,
          STDDEV(eff.fuel_efficiency_per_km) AS std_dev_fuel_efficiency_per_km
          FROM
          passenger co2
          INNER JOIN effciency eff
          ON co2.code = eff.code AND co2.year = eff.year
          GROUP BY
          co2.entity, co2.year;
           ") %>%
  ggplot(aes(x = avg_fuel_efficiency_per_km, y = avg_co2_per_km)) +
  geom_point() +
  facet_wrap(~country) +
  labs(title = "CO2 Emissions vs. Fuel Efficiency by Country and Year",
       x = "Fuel Efficiency (km/L)",
       y = "CO2 Emissions (g/km)")
```

Visualize the data: You can create various plots to visualize the data, such as scatter plots, line graphs, and box plots. For example, you could create a scatter plot of CO2 emissions per km against fuel efficiency per km for each country and year to see if there is a correlation between the two variables.

```{r}
dbGetQuery(con,"
SELECT
  co2.entity AS country,
  co2.year,
  AVG(co2.co2_per_km) AS avg_co2_per_km,
  STDDEV(co2.co2_per_km) AS std_dev_co2_per_km,
  AVG(eff.fuel_efficiency_per_km) AS avg_fuel_efficiency_per_km,
  STDDEV(eff.fuel_efficiency_per_km) AS std_dev_fuel_efficiency_per_km
FROM
  passenger co2
  INNER JOIN effciency eff
  ON co2.code = eff.code AND co2.year = eff.year
GROUP BY
  co2.entity, co2.year;")

```

```{r}
fuel_efficiency %>% ggplot(aes(x = fuel_efficiency_per_km, y = co2_per_km, color = country)) +
  geom_point() +
  facet_wrap(~year) +
  labs(title = "CO2 Emissions vs. Fuel Efficiency by Country and Year",
       x = "Fuel Efficiency (km/L)",
       y = "CO2 Emissions (g/km)")
```

Sure! Here's an example SQL code to calculate the z-scores for the co2_per_km column and identify any observations with a z-score greater than a certain threshold (in this case, 3):

```{r}
dbGetQuery(con,"
           SELECT
  entity,
  year,
  co2_per_km, z_score
FROM
  (SELECT
     entity,
     year,
     co2_per_km,
     ABS((co2_per_km - AVG(co2_per_km) OVER ()) / STDDEV(co2_per_km) OVER ()) AS z_score
   FROM
     passenger) subquery
WHERE
  z_score < 3
ORDER BY
  z_score DESC;
           ") %>%
  
  ggplot(aes(x = year, y = co2_per_km, color = z_score > 3)) +
  geom_point() +
  facet_wrap(~entity, scales = "free") +
  scale_color_manual(values = c("FALSE" = "blue", "TRUE" = "red"), name = "Outlier") +
  labs(x = "Year", y = "CO2 per km", title = "Identifying outliers using z-scores")

```

Sure, here's an example SQL query that compares the average CO2 emissions per km and fuel efficiency per km across different countries and years:


```{r}
dbGetQuery(con,"
           SELECT
  co2.entity AS country,
  co2.year,
  AVG(co2.co2_per_km) AS avg_co2,
  AVG(eff.fuel_efficiency_per_km) AS avg_fuel_efficiency
FROM
  passenger co2
  INNER JOIN effciency eff
    ON co2.code = eff.code AND co2.year = eff.year
GROUP BY
  co2.entity,
  co2.year
ORDER BY
  co2.entity,
  co2.year;")
```

```{r}
dbGetQuery(con,"
SELECT co2.entity AS country, co2.year, AVG(co2.co2_per_km) AS avg_co2, AVG(eff.fuel_efficiency_per_km) AS avg_fuel_efficiency
          FROM passenger co2
          INNER JOIN effciency eff
          ON co2.code = eff.code AND co2.year = eff.year
          GROUP BY co2.entity, co2.year
          ORDER BY co2.entity, co2.year;") %>%
# Create the ggplot
ggplot(aes(x = country, y = avg_fuel_efficiency, fill = year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Country", y = "Average fuel efficiency per km", fill = "Year", title = "Comparing average CO2 emissions and fuel efficiency by country and year") + coord_flip()
```

